Epistemic Semantics in Guarded String Models

### 2020-06-16 tex/draft.tex
Figure with term syntax in draft.tex

@@

\pagebreak
\section{Syntax-semantics interface}
An architecture of interpretation by transtlation is employed, where English
sentences are mapped to terms in the logical language (--) via an interpreted
grammar, and these terms are in turn interpreted as propositions (sets of possible
worlds). For the latter, there are options of translation into the finite state
calculus in order to represent propositions as finite state machines (Section 3), and
representation in Haskell via lazy lists of guarded strings (Section 4).  The grammar
is a semantically interpreted multi-modal categorial grammar, consisting of a lexicon
of words, their categorial types, and interpretations in a logical lambda language.
\ex{phenomena1} lists that are covered.

\begin{example}   \label{phenomena1}
{\em Phenomena}

\vspace{1mm}
\hspace{-10mm}\begin{tabular}[t]{l|l}
  Basic statives & It's heads. \\
  & It's tails.\\ \hline
That-complement & Amy knows \\ & \hspace{2mm} that it's heads. \\ \hline
Wh-complement & Amy knows \\ & \hspace{2mm} whether its heads. \\ \hline
Negation & Bob doesn't know \\ & \hspace{2mm} that it isn't heads. \\ \hline 
Tensed and base &  Bob knows \\
verbal forms & \hspace{2mm} that it's heads.      \\ 
\hspace{1mm} & Bob doesn't know \\ & \hspace{2mm} that it's heads.\\ \hline
Sentence conjunction & It's heads and Bob \\ & \hspace{2mm} doesn't know \\ & \hspace{2mm} that it's heads \\ \hline
Predicate conjunction & Bob knows that \\ & Amy knows whether \\ & it's heads and
\\ & doesn't know that \\ & Amy knows that \\ & it's heads.
\end{tabular}    
\end{example}

>>>>

This scheme provides for an interpretation of the language of Epik
terms that is defined in Figure 4.  An Epik specifications such as
Figure 1 is translated to a straight-line program of the finite state
calculus that defines constants and functions, including the ones from
Figures 2 and 3.  In Xfst or Foma, which are interpreters for the
finite state calculus, the program is read, and then propositional
terms can be mapped to finite state machines that represent sets of
guarded strings, interpreted as propositions.

@@

\section{Syntax-semantics interface}

\begin{figure}
\begin{tabular}{lll}
agents & $a,b...$ & $a {:}{:}{=} \, \tuple{\mbox{atomic agents}}$ 
\end{tabular}

\begin{tabular}{lll}
tests & $T,S...$ & $T{:}{:}{=} \, \begin{array}[t]{l}\tuple{\mbox{atomic tests}}\,|\,0\,|\,1\,|\\ T{+}S\;|\;T\,{;}\,S\;|\;!T\end{array}$
\end{tabular}

\begin{tabular}{lll}
propositions & $p,q...$ & \(p{:}{:}{=} \,
\begin{array}[t]{l}\tuple{\mbox{atomic events}}\,| \\
T \,|\, 
p{+}q \,|\,
p\mbox{\&}q \,|\, \\
p{;}q \;|\;
p{*} \;|\; {\sim}p \;|\\
{\diamond} a p
\end{array}\)
\end{tabular}

\caption{The language of Epik terms. Propositional terms are interpreted
as sets of guarded strings.}
\end{figure}

An architecture of interpretation by transtlation is employed, where English
sentences are mapped to terms in the logical language (--) via an interpreted
grammar, and these terms are in turn interpreted as propositions (sets of possible
worlds). For the latter, there are options of translation into the finite state
calculus in order to represent propositions as finite state machines (Section 3), and
representation in Haskell via lazy lists of guarded strings (Section 4).  The grammar
is a semantically interpreted multi-modal categorial grammar, consisting of a lexicon
of words, their categorial types, and interpretations in a logical lambda language.
\ex{phenomena1} lists that are covered.

\begin{figure}  
\begin{tabular}[t]{l|l}
  Basic statives & It's heads. \\
  & It's tails.\\ \hline
That-complement & Amy knows \\ & \hspace{2mm} that it's heads. \\ \hline
Wh-complement & Amy knows \\ & \hspace{2mm} whether its heads. \\ \hline
Negation & Bob doesn't know \\ & \hspace{2mm} that it isn't heads. \\ \hline 
Tensed and base &  Bob knows \\
verbal forms & \hspace{2mm} that it's heads.      \\ 
\hspace{1mm} & Bob doesn't know \\ & \hspace{2mm} that it's heads.\\ \hline
Sentence conjunction & It's heads and Bob \\ & \hspace{2mm} doesn't know \\ & \hspace{2mm} that it's heads \\ \hline
Predicate conjunction & Bob knows that \\ & Amy knows whether \\ & it's heads and
\\ & doesn't know that \\ & Amy knows that \\ & it's heads.
\end{tabular}
\caption{Phenomena covered in the English grammar fragment.}
\end{figure}
